---
title: "rMEM"
author: "James T Morris, James R Holmquist, Kathe Todd-Brown"
date: "10/2/2019"
output: html_document
---

The marsh equilibrium model is a concept developed by Morris et al over several iterations. Different versions have in common that as sea-levels rise marsh surfaces dynamically respond to increasing inundation. Organic 

MEM is a 0 dimensional model, meaning it treats spatial dynamics as a single point with no input from or output to other connected patches of the marsh. This makes it relatively simple and tractable.  

Different versions of MEM have these elements in common.

This is a full-blown implementation of MEM. It can be used for hind-casting long-term carbon burial rates, and or hindcasting forecasting carbon removal and net-elevation change. The MEM inputs 

New optionalities include.    
  *  More optionalities for inputing biomass parabolas.  
  *  Option to incorporate sediment deposition accross 3 different tidal ranges (high, higher high, and spring tides).  
  *  More flexibility for inputing custom sea-level change, suspended sediment concentraton scenarios.  
  *  Custom specification of depth intervals for soil cores.  
  *  Options to create animations of soil profile formation, sea-level and surface elevation change, and biomass and elevation response.  

## Load R CTM
The first thing we will do is load the cohort theory model as a dependency.

```{r setup, include=TRUE}
# devtools::install_github("https://github.com/tilbud/rCTM")
require(rCTM)
require(tidyverse)
require(gridExtra)
#source("../R/predictedBiomass.R")
#source("../R/zToZstar.R")
#source("../R/zStarToZ.R")

```

## Define Some Inputs

``` {r define inputs}

# Settings for the functions we're going to run
# Inputs
startYear <- 2015 # yearStart
rslr0 <- 0.3 # rslr at time 1 (cm/yr)
rslrTotal <- 100 # cm of total relative seal level rise
initElv <- 21.9 # eInit # cm relative to NAVD88
MSL <- 7.4 # tidal datums, cm relative to NAVD88
MHW <- 16.9
MHHW <- 25.4
MHHWS <- 31.2
ssc <- 30 # milligrams per liter /  grams per m3
ssc <- ssc / 1000000 # convert to grams per cubic cm
lunarNodalAmp <- 5 # cm

# Parameters
bMax <- 2500 # grams per m-2
bMax <- bMax/10000 # grams per cm-2
zVegMin <- -24.7 # cm relative to datum
zVegMax <- 44.4
zVegPeak <- 22.1
plantElevationType = "orthometric"
rootToShoot <- 2 # root to shoot ratio
rootTurnover <- 0.5 # n / year
rootDepthMax <- 30 # cm surface
omDecayRate <- 0.8
recalcitrantFrac <- 0.2
settlingVelocity <- 2.8 # n clearences / tide cycle

```

## Build Biomass Parabolas Function

This function takes relative elevation (Z) as an input. It also has parameters descriping parabolic relationship between Z\* and above ground biomass production (grams dry mass m^-2^). It has three manditory parameters, maximum biomass, and minimum and maximum growing elevations.

``` {r parabolicBiomass}

agbFromElv <- predictedBiomass(z=initElv, bMax = bMax, zVegMax = zVegMax,
                               zVegMin = zVegMin, zVegPeak = zVegPeak)

(agbFromElv)

print(agbFromElv)

```

## Create Helper Scripts for Converting Too and From Dimensionless Tidal Elevation

In the biomass functions the units for the elevation can be any thing as long as they are all referenced to the same system. We prefer elevation relative to mean sea level (MSL), normalized to the daily high tide range, mean high water (MHW) This is sometimes referred to as a dimensionless tidal elevation (Z\*). Using Z\* makes biomass curves more comparable accross sites and times, accross which MSL and tidal ranges can vary.  

$$Z^*={E-MSL\over{MHW-MSL}}$$

If you use the most basic settings the function will assume median growing biomass is exactly in between min and max. If you specify a median the function will create a piecewise, flexible parabola.

Here are two scripts for quickly converting to and from dimensionless tidal elevation.

```{r zStarFunctions}
# Target Elevation
targetElevation<- seq((zVegMin-abs(zVegMin*.1)), (zVegMax+abs(zVegMax*.1)), 
                      by = ((zVegMax-zVegMin)/100) )

# Convert elevation to dimensionless tidal elvation
zStarExample <- zToZstar(z=targetElevation, MSL=MSL, MHW=MHW)

ggplot(data=data.frame(z=targetElevation, zStar = zStarExample), aes(x=z, y=zStar)) +
  geom_line() +
  geom_vline(data=(z=data.frame(z=c(MHW, MSL), 
                                datum=c("MHW", "MSL"))), 
             aes(xintercept = z, lty=datum), color = "blue") +
  geom_hline(data=(z=data.frame(z=c(1, 0), 
                                datum=c("MHW", "MSL"))), 
             aes(yintercept = z, lty=datum), color = "red")

# Convert dimensionless tidal elevation back to regular elevation
zReconstructExample <- zStarToZ(zStar = zStarExample, MSL = MSL, MHW = MHW)

# See they're the same
head(targetElevation)
head(zReconstructExample)

```


``` {r exampleParabolas, fig.height = 3.54, fig.width=7}

# targetElevations <- seq(min(zVegMin, MSL-MHHWS), max(zVegMax, MHHWS), by = 4)

displayAGB <- data.frame(elevation=targetElevation) %>%
  mutate(`no peak evl.`= predictedBiomass(z=targetElevation, bMax=bMax, zVegMax=zVegMax, 
                                          zVegMin=zVegMin, zVegPeak=NA),
         `w/ peak elv.`= predictedBiomass(z=targetElevation, bMax=bMax, zVegMax=zVegMax, 
                                          zVegMin=zVegMin, zVegPeak=zVegPeak)) %>%
  gather(value="agb", key='parabolaType', -elevation)


# and plot
ggplot(data=displayAGB, aes(x=elevation, y=agb)) +
  geom_line(color="red", lwd=1.25) +
  facet_wrap(.~parabolaType) +
  ylab(expression(paste("Above Ground Biomass (g cm"^"-2", ")", sep="")))

```

## Build Sedimentation Functions

Sedimentation in MEM is a function of suspended sediment concentration, and three properties of elevation relative to tidal range, the number of tides a year, the mean flood depth, and the mean flood time per tidal cycle. 

The original version of MEM makes some simplistic assumptions about n tides, and flood time. Original MEM assumes that no sedimentation occurs above the MHW line.

```{r simple flood time}

avaliableSediment <- function(floodPct, ssc, settlingVelocity, k=1) {
  
  availableSSC <- ifelse(floodPct < 1/settlingVelocity, # if the sediment column IS NOT able to clear
                         # available ssc is total possible caputre 
                         ssc * floodPct/(1/settlingVelocity) * k, 
                         # if the sediment column IS able to clear
                         ssc * k)
  
  return(availableSSC)
}

deliveredSedimentSimple <- function(z, ssc, MSL, MHW, MLW = MSL-MHW, settlingVelocity, 
                                    k=1, nTides=704) {
  
  # Flood depth in meters is the same as water volume when in m and assuming a 1m2 area of interest
  floodDepth <- ifelse(z<=MHW, (MHW-z)*0.5, 0)
  
  # Mean flood percent is a line, relative position of elevation between MHW and MLW.
  floodPct <- ifelse(z >= MHW, 0, ifelse(z <= MLW, 1, (MHW-z)/(MHW-MLW)))
  
  availableSSC <- avaliableSediment(ssc=ssc, floodPct=floodPct, settlingVelocity=settlingVelocity, k=k)
  
  # Mean flood depth is depth below MHW. If above MHW the marsh does not flood.
  # delivered sediment is the ssc available to the surface (g/m3) multiplied by 
  # the cumulative water volume (cube of water) and number of times the cube passes over the marsh. 
  deliveredSSC <- availableSSC * floodDepth * nTides
  
  return(deliveredSSC)
}

```

Here's an example.

```{r simple sediment example, fig.width=3.54, fig.height=3.54}

highTidesPerYear <- 352.657
higherHighTidesPerYear <- 352.657 - 24.720
springTidesPerYear <- 24.720

weightedMHW <- sum(MHHWS*springTidesPerYear + MHHW*higherHighTidesPerYear + MHW*highTidesPerYear) /
  sum(highTidesPerYear, higherHighTidesPerYear, springTidesPerYear)

deliveredSedimentExample <- mapply(deliveredSedimentSimple, z=targetElevation, MSL=MSL, MHW=weightedMHW,
                                   settlingVelocity = settlingVelocity, ssc=ssc)

simpleSedimentPlotData <- data.frame(elevation=targetElevation,
                                     deliveredSediment=deliveredSedimentExample)

ggplot(data=simpleSedimentPlotData, aes(x=elevation, y=deliveredSediment)) +
  geom_line(lwd=1.25) +
  ylab(expression(paste("Delivered Sediment (g cm"^"-2",")", sep="")))

```

_Question? Is there a lower limit on sediment delivery to the marsh in MEM? Is it MLW or the lower vegetation line?_

```{r flood time accross datums }

floodTimeFromDatum <- function(z, datumHigh, datumLow, ...) {
  
  # If elevation is above the tidal range indation time is 0
  datumHigh <- ifelse(z>=datumHigh, z, datumHigh)
  
  # If elevation is below inundation time is a full tidal cycle
  datumLow <- ifelse(z<=datumLow, z, datumLow)
  
  # Rising time over cell = 6.21 (A/pi - 1)	
  # where A = 2* pi - cos-1 [2 (height of cell – MLW) / (MHW – MLW) - 1] radians
  A1 <- 2 * pi - acos(2 * (z-datumLow) / (datumHigh-datumLow) - 1)
  risingTime <- 6.21 * (A1/pi - 1)
  
  # Falling time over cell = 6.21 (A/pi - 1) where
  # A = 2* - cos-1 [2 (height of cell – MHW) / (MLW – MHW) - 1] radians
  A2 <- 2 * pi - acos(2 * (z-datumHigh) / (datumLow-datumHigh) - 1)
  fallingTime <- 6.21 * (A2/pi - 1)

  # If between inundation time = abs (time rising - 6.21) + time falling
  inundationTime <- abs(risingTime - 6.21) + fallingTime

  return(inundationTime)
}

deliveredSediment3TidalCycle <- function(z, ssc, MSL, MHW, MHHW, MHHWS,
                                        MLW=MSL-MHW, MLLW=MSL-MHHW, MLLWS=MSL-MHHWS,
                                        settlingVelocity, k=1) {
  
  # Define Constants
  highTidesPerYear <- 352.657
  higherHighTidesPerYear <- 352.657 - 24.720
  springTidesPerYear <- 24.720
  hoursInTidalCycle <- 12.42
  
  # Create a data frame operation so we can use tidy functions to speed this up
  tidalCycles <- data.frame(datumHigh = c(MHW, MHHW, MHHWS), 
                            datumLow = c(MLW, MLLW, MLLWS),
                            nTides = c(highTidesPerYear, higherHighTidesPerYear, springTidesPerYear))
  
  tidalCycles <- tidalCycles %>%
    # Set tidal propoerties to 0 if surface is above tidal range in each case
    dplyr::mutate(nTides = ifelse(z>datumHigh, 0, nTides), # number of tides
           tidalHeight = ifelse(z>datumHigh, 0, (datumHigh-z)*0.5), # Tidal height relative to surface
           floodTime = ifelse(z>datumHigh, 0, floodTimeFromDatum(z=z, # Call flood time function.
                                                                 datumHigh=datumHigh, 
                                                                 datumLow=datumLow)), 
           floodPct = ifelse(z>datumHigh, 0, floodTime / hoursInTidalCycle), # Convert hours to fraction
           # Call available sediment function
           availableSSC = avaliableSediment(floodPct=floodPct, ssc=ssc, 
                                            settlingVelocity=settlingVelocity, k=k),
           deliveredSSC = availableSSC * nTides * tidalHeight) # Calculated delivered SSC
  
  # Sum delivered SSC accross tidal cycles
  totalDeliveredSSC <- sum(tidalCycles$deliveredSSC) 
  
  return(totalDeliveredSSC)
}

# Function decides which sediment module to use based on available inputs
deliverSedimentFlexibly <- function(z, ssc, MSL, MHW, MHHW=NA, MHHWS=NA, settlingVelocity) {
  # If the user does not include detailed tidal datums
  if (any(is.na(c(MHHW, MHHWS)))) {
    # run the simple SSC module
    deliveredSSC <- deliveredSedimentSimple(z=z, ssc=ssc, MSL=MSL, MHW=MHW, settlingVelocity=settlingVelocity)
  } else {
    # If they do include them, run the 3 tide stage module
    deliveredSSC <- deliveredSediment3TidalCycle(z=z, ssc=ssc, MSL=MSL, MHW=MHW, settlingVelocity=settlingVelocity,
                                                 MHHW=MHHW, MHHWS=MHHWS)
  }
  return(deliveredSSC)
}

```

Let's see how the simple vs two tidal cycle modules compare. 

```{r example of simple vs complex sedimentation modules, warning=FALSE, fig.height=3.54, fig.width=7}

deliveredSedimentExample2 <- mapply(deliveredSediment3TidalCycle, z=targetElevation, ssc=ssc, MSL=MSL,
                                    MHW=MHW, MHHW=MHHW, MHHWS=MHHWS, settlingVelocity=settlingVelocity)

threeTideSedimentPlotData <- data.frame(elevation=targetElevation,
                                        deliveredSediment = deliveredSedimentExample2,
                                        moduleType = rep("3 tide", length(targetElevation)))

simpleSedimentPlotData$moduleType <- rep("simple ", nrow(simpleSedimentPlotData))

sedimentComparisonPlot <- threeTideSedimentPlotData %>%  bind_rows(simpleSedimentPlotData)

totalRange <- ggplot(data=sedimentComparisonPlot, aes(x=elevation, y=deliveredSediment)) +
  geom_line(lwd=1.25, aes(color=moduleType), alpha=0.6) +
  ylab(expression(paste("Delivered Sediment (g cm"^"-2",")", sep=""))) +
  ggtitle("Total Tidal Range") +
  geom_vline(data=data.frame(elevation=c(MSL, MHW, MHHW, MHHWS),
                             datum=c("MSL", "MHW", "MHHW", "MHHWS")), aes(xintercept=elevation,
                                                                          lty=datum))

sedimentComparisonPlotZoomedIn <- filter(sedimentComparisonPlot, elevation>MHW)

# Zoom In
zoomedIn <- ggplot(data=sedimentComparisonPlotZoomedIn, aes(x=elevation, y=deliveredSediment)) +
  geom_line(lwd=1.25, aes(color=moduleType), alpha=0.6) +
  ylab(expression(paste("Delivered Sediment (g cm"^"-2",")", sep=""))) +
  ggtitle("Zoomed in Above MHW")  +
  geom_vline(data=data.frame(elevation=c(MHW, MHHW, MHHWS),
                             datum=c("MHW", "MHHW", "MHHWS")), 
             aes(xintercept=elevation, lty=datum))

grid.arrange(totalRange,zoomedIn, ncol=2)

```

## Build Mean-Sea Level Curve

``` {r build sea-level curve}

buildScenarioCurve <- function(startYear, endYear=startYear+99, MSL, rslr0=0.3, rslrTotal=100, ssc) {
  
  # Create a sequence of the number of years
  years <- startYear:endYear
  nYearsOfSim <- length(years) # Put this in args. Replace this with yearEnd.
  
  scenario <- data.frame(index = 1:nYearsOfSim, years = years,
                         MSL = rep(NA, nYearsOfSim),
                         ssc = rep(NA, nYearsOfSim))
  
  # Build the Mean Sea Level Scenario
  # If the input only specifies an initial Mean Sea Level at time = 0 ... 
  if (length(MSL) == 1) {
    # ... create a sea-level rise scenario based on total sea-level rise and inital relative sea-level rise rate
    # Probably could use a stop here in case there are invalid entries of rslr0 and slrTotal
    scenario$MSL[1] <- MSL
    
    # IPCC, 2013 and Sweet et al., 2017
    # MSL(t) = MSL(0) + At + Bt^2 - where t is years from time zero
    # A=[MSL(1)-MSL(0)] – B  
    # Where [MSL(1)-MSL(0)] is the SLR at the start as determined from NOAA trend analyses 
    # Coefficient B is the acceleration term 
    # B = {[MSL(T)-MSL(0)] / T - [MSL(1)-MSL0] } / (T-1) where T - length of the simulation
    
    B <- ((rslrTotal)/nYearsOfSim - rslr0) / (nYearsOfSim-1)
    A <- rslr0 - B
    
    scenario$MSL[2:nYearsOfSim] <- scenario$MSL[1] + A*scenario$index[2:nYearsOfSim] + B*scenario$index[2:nYearsOfSim]^2
    
  } else if (length(MSL) == length(years)) {
    
    # If the user enters in a vector of MSL that is equal to the number of years in the simulation.
    scenario$MSL <- MSL
    
  } else {
    stop("RSLR input is incorrect. Either enter a value at for the beginning and ending of the scenario, or a vector of RSLR one for each year of the scenario.")
  }

  # Add suspended sediment concentration
  if (length(ssc) == 1) {
    # If ssc is a single value.
    scenario$ssc <- rep(ssc, nYearsOfSim)
  } else if (length(ssc) == nYearsOfSim) {
    # If ssc is a vector in equal length to the number of years of the simulation.
    scenario$ssc <- ssc
  } else {
    stop("SSC input is incorrect. Either enter a single average value, or a vector equal in length to the number of years in the scenario.")
  }
  return(scenario)
  # This is a start. We could also allow flexibility to change biomass and sediment addition functions later on to incorporate things like conversions of emergent to forested wetland.
}

```

## Build MHW Lines

``` {r build tidal range scenario}

predictLunarNodalCyle <- function(year, floodElv, MSL0, MSL, lunarNodalAmp) {
  # Build MHW lines based on MSL, long-term tidal range and lunar nodal amplitude
  MHW <- MSL + (floodElv-MSL0) + (lunarNodalAmp * (sin(2*pi*(year-1983)/18.61)))
  return(MHW)
}

buildHighTideScenario <- function(scenarioCurve, MHW0, MHHW0, MHHWS0, lunarNodalAmp) {
  
  # In create a MHW and add it to the scenario
  scenarioCurve$MHW <- predictLunarNodalCyle(year = scenarioCurve$years, MSL= scenarioCurve$MSL, 
                                             MSL0 = scenarioCurve$MSL[1], floodElv=MHW0, 
                                             lunarNodalAmp=lunarNodalAmp)
  
  # If MHHW and MHHWS are arguements add them to the scenario table too
  if (!missing(MHHW0) & !missing(MHHWS0)) {
    scenarioCurve$MHHW <- predictLunarNodalCyle(year = scenarioCurve$years, MSL=scenarioCurve$MSL,
                                                MSL0 = scenarioCurve$MSL[1], floodElv=MHHW0, 
                                                lunarNodalAmp=lunarNodalAmp)
    scenarioCurve$MHHWS <- predictLunarNodalCyle(year = scenarioCurve$years, MSL= scenarioCurve$MSL, 
                                                 MSL0 = scenarioCurve$MSL[1], floodElv=MHHWS0, 
                                                 lunarNodalAmp=lunarNodalAmp)
  }
  return(scenarioCurve)
}
```

### High Tide Scenario Curve Example
```{r build scenario}
exScenarioCurve <- buildScenarioCurve(startYear = 2015, MSL = MSL, rslr0=0.3, 
                                      rslrTotal = 100, ssc=30)
head(exScenarioCurve)

exScenarioCurve <- buildHighTideScenario(scenarioCurve=exScenarioCurve, MHW0=MHW, MHHW0=MHHW, 
                                         MHHWS0=MHHWS, lunarNodalAmp = lunarNodalAmp)
head(exScenarioCurve)

ggplot(data=exScenarioCurve) + 
  geom_ribbon(aes(x=years, ymax=MHW, ymin=MSL-(MHW-MSL)), alpha=0.6) +
  geom_line(aes(x=years, y=MSL), color="black") +
  geom_point(aes(x=years, y=MSL), color="black", pch=16) +
  ylab("Water Level (cm NAVD88)") +
  ggtitle("Sea-Level Rise and Tidal Amplitude Scenario.")

highTideScenarios <- exScenarioCurve %>%
  select(-ssc, -index) %>%
  gather(value="elevation", key="high tide", -years, -MSL) %>%
  arrange()

ggplot(data=exScenarioCurve) + 
  geom_line(aes(x=years, y=MSL), color="black") +
  geom_point(aes(x=years, y=MSL), color="black", pch=16) +
  geom_ribbon(data = highTideScenarios, aes(x=years, ymax=elevation, ymin=MSL-(elevation-MSL), 
                                            color=`high tide`),fill=NA) +
  ylab("Water Level (cm NAVD88)") +
  ggtitle("Sea-Level Rise and Multi-Tide Amplitude Scenario.") +
  theme(legend.title = element_blank())

```


## Run all functions together

```{r linked MEM CTM } 

trimCohorts <- function(cohorts) {
  cohorts <- cohorts %>%
    dplyr::filter(cumCohortVol!=0)
  return(cohorts)
}

runMemWithCohorts <- function(startYear, endYear=startYear+99, rslrT1, rslrTotal, initElv,
                              MSL, MHW, MHHW=NA, MHHWS=NA, ssc, lunarNodalAmp,
                              bMax, zVegMin, zVegMax, zVegPeak, plantElevationType,
                              rootToShoot, rootTurnover, rootDepthMax, shape="linear",
                              omDecayRate, recalcitrantFrac, settlingVelocity,
                              omPackingDensity=0.085, mineralPackingDensity=1.99,
                              rootPackingDensity=omPackingDensity,
                              coreYear=NA, coreDepth=100, coreMaxs=1:100, coreMins=0:99,
                              ...) {
  
  # Run checks on input and trigger warnings if units seem off
  # Need to develop this
  
  # Build scenario curve
  scenario <- buildScenarioCurve(startYear=startYear, endYear=endYear, MSL=MSL, 
                                 rslr0=rslr0, rslrTotal=rslrTotal, ssc=ssc)
  
  # add high tides
  scenario <- buildHighTideScenario(scenario, MHW0=MHW, MHHW0=MHHW, MHHWS0=MHHWS, 
                                    lunarNodalAmp = lunarNodalAmp)
  
  # Add blank colums for attributes we will add later
  scenario$surfaceElevation <- rep(NA, nrow(scenario))
  scenario$biomass <- rep(NA, nrow(scenario))
  scenario$mineral <- rep(NA, nrow(scenario))
  scenario$cumulativeTotalOm <- rep(NA, nrow(scenario)) 
  
  # Convert dimensionless plant growing elevations to real growing elevations
  if (! plantElevationType %in% c("dimensionless", "zStar", "Z*", "zstar")) {
    zStarVegMin <- zToZstar(zVegMin, MHW, MSL[1])
    zStarVegMax <- zToZstar(zVegMax, MHW, MSL[1])
    zStarVegPeak <- zToZstar(zVegPeak, MHW, MSL[1])
  } else {
    zStarVegMin <- zVegMin
    zStarVegMax <- zVegMax
    zStarVegPeak <- zVegPeak
  }

  # Set initial conditions
  # Calculate initial z star
  initElvStar <- zToZstar(z=initElv, MSL=scenario$MSL[1], MHW=scenario$MHW[1])
  
  # Initial Above Ground Biomass
  initAgb <- predictedBiomass(z=initElvStar, bMax = bMax, zVegMax = zStarVegMax, 
                              zVegMin = zStarVegMin, zVegPeak = zStarVegPeak)
  
  # Initial Below Ground Biomass
  initBgb <- initAgb * rootToShoot
  
  # Initial Sediment
  initSediment <- deliverSedimentFlexibly(z=initElv, ssc=scenario$ssc[1], 
                                          MSL=scenario$MSL[1], MHW=scenario$MHW[1], 
                                          MHHW = scenario$MHHW[1], MHHWS = scenario$MHHWS[1],
                                          settlingVelocity=settlingVelocity)
  
  depositInitSediment <- function(...){return(initSediment)}
  
  # Run initial conditions to equilibrium
  cohorts <- runToEquilibrium(totalRootMass_per_area=initBgb, rootDepthMax=rootDepthMax,
                              rootTurnover=rootTurnover, omDecayRate = list(fast=omDecayRate, slow=0),
                              rootOmFrac=list(fast=1-recalcitrantFrac, slow=recalcitrantFrac),
                              packing=list(organic=omPackingDensity, mineral=mineralPackingDensity),
                              rootDensity=rootPackingDensity, shape=shape, 
                              mineralInput_g_per_yr.fn = depositInitSediment)
  
  # Add initial conditions to annual time step tracker
  scenario$surfaceElevation[1] <- initElv
  scenario$mineral[1] <- initSediment
  scenario$biomass[1] <- initBgb
  scenario$cumulativeTotalOm[1] <- sum(cohorts$fast_OM, cohorts$root_mass, cohorts$slow_OM, na.rm=T) 
  
  # Trim NA rows from eq cohorts
  cohorts <- trimCohorts(cohorts)
  cohorts$year <- rep(scenario$years[1], nrow(cohorts))
  
  # Preallocate memory for cohort tracking
  nInitialCohorts <- nrow(cohorts)
  nScenarioYears <- nrow(scenario)
  initCohortRows <- nInitialCohorts * nScenarioYears
  newCohortRows <- sum(1:nScenarioYears)
  totalRows <- initCohortRows+newCohortRows
  
  trackCohorts <- data.frame(age=rep(rep(NA, totalRows)),
                             fast_OM=rep(NA, totalRows), slow_OM=rep(NA, totalRows), 
                             mineral=rep(NA, totalRows), root_mass=rep(NA, totalRows), 
                             layer_top=rep(NA, totalRows), layer_bottom=rep(NA, totalRows),
                             cumCohortVol=rep(NA, totalRows),
                             year=rep(NA, totalRows))
  
  # add initial set of cohorts
  trackCohorts[0:nInitialCohorts,] <- cohorts
  
  # create variables to keep track of cohorts added to the full cohort tracker
  cohortsNewRowMin <- length(trackCohorts$age[!is.na(trackCohorts$age)]) + 1
  
  # Calculate the unmoving bottom of the profile as a consistent reference datum
  profileBottomElv <- initElv - max(cohorts$layer_bottom)
  
  # Fourth, add one cohort for each year in the scenario
  # Iterate through scenario table
  for (i in 2:nrow(scenario)) {
    
    # Calculate surface elevation relative to datum
    surfaceElvZStar <- zToZstar(z=scenario$surfaceElevation[i-1], MHW=scenario$MHW[i], MSL=scenario$MSL[i])
    
    # Caluclate dynamic root biomass
    dynamicAgb <- predictedBiomass(z=surfaceElvZStar, bMax = bMax, zVegMax = zStarVegMax, zVegMin = zStarVegMin, zVegPeak = zStarVegPeak)
  
  # Initial Below Ground Biomass
  dynamicBgb <- dynamicAgb * rootToShoot
  
  dynamicMineralPool <- deliverSedimentFlexibly(z=scenario$surfaceElevation[i-1], ssc=scenario$ssc[i], 
                                          MSL=scenario$MSL[i], MHW=scenario$MHW[i], MHHW = scenario$MHHW[i], 
                                          MHHWS = scenario$MHHWS[i], settlingVelocity=settlingVelocity)
  
  depositMineralPool <- function(...){return(dynamicMineralPool)}
    
  # Calculate dynamic sediment deliver
  cohorts <- addCohort(cohorts, totalRootMass_per_area=dynamicBgb, rootDepthMax=rootDepthMax, 
                       rootTurnover = rootTurnover, omDecayRate = list(fast=omDecayRate, slow=0),
                       rootOmFrac=list(fast=1-recalcitrantFrac, slow=recalcitrantFrac),
                       packing=list(organic=omPackingDensity, mineral=mineralPackingDensity), 
                                       rootDensity=rootPackingDensity, shape=shape,
                       mineralInput_g_per_yr.fn = depositMineralPool)
    
    cohorts <- trimCohorts(cohorts)
    cohorts$year <- rep(scenario$years[i], nrow(cohorts))
    
    scenario$surfaceElevation[i] <- profileBottomElv + max(cohorts$layer_bottom, na.rm=T)
    scenario$biomass[i] <- dynamicBgb
    scenario$mineral[i] <- dynamicMineralPool
    scenario$cumulativeTotalOm[i] <- sum(cohorts$fast_OM, cohorts$root_mass, cohorts$slow_OM, na.rm=T) 
    
    # add initial set of cohorts
    cohortsNewRowMax <- cohortsNewRowMin + nrow(cohorts)-1
    trackCohorts[cohortsNewRowMin:cohortsNewRowMax,] <- cohorts
  
    # create variables to keep track of cohorts added to the full cohort tracker
    cohortsNewRowMin <- length(trackCohorts$age[!is.na(trackCohorts$age)]) + 1
  }
  
  # Return annual time steps and full set of cohorts
  outputsList <- list(annualTimeSteps = scenario, cohorts = trackCohorts)
  
  return(outputsList)
}

```

``` {r graph MEM-CTM output}
# First we'll profile the script
start_time <- Sys.time()
memCohortExample <- runMemWithCohorts(startYear=2015, rslrT1=0.3, rslrTotal=100,
                                      initElv=21.9, MSL=7.4, MHW=16.9, MHHW=25.4, MHHWS=31.2, 
                                      ssc=3e-05, lunarNodalAmp=5, bMax=0.25, 
                                      zVegMin=-24.7, zVegMax=44.4, zVegPeak=22.1,
                                      plantElevationType="orthometric", rootToShoot=2,
                                      rootTurnover=0.5, rootDepthMax=30, omDecayRate=0.8,
                                      recalcitrantFrac=0.2, settlingVelocity=2.8
                                      )
print(Sys.time()-start_time)

scenario <- memCohortExample$annualTimeSteps
cohorts <- memCohortExample$cohorts

scenarioNetElevationChange <- scenario %>%
  select(years, MSL, surfaceElevation) %>%
  gather(key="Elevation Type", value="Elevation", -years)

plot1 <- ggplot(data=scenarioNetElevationChange, aes(x=years, y=Elevation)) +
  geom_line(aes(color=`Elevation Type`)) +
  ggtitle("Elevation vs Sea-Level") +
  theme(legend.title = element_blank(),
        legend.position="top")

plot2 <- ggplot(data=scenario, aes(x=years, y=biomass)) +
  geom_line() +
  ggtitle("Belowground Biomass Over Time")

plot3 <- ggplot(data=scenario, aes(x=years, y=mineral)) +
  geom_line() +
  ggtitle("Mineral Deposition Over Time")

plot4 <- ggplot(data=scenario, aes(x=years, y=cumulativeTotalOm)) +
  geom_line() +
  ggtitle("Cumulative OM deposited over time")

grid.arrange(plot1, plot2, plot3, plot4, ncol=2)

```

__Here is an example of something you can't do with simple MEM.__

``` {r fullBlownMem exampl}
scenarioFlux <- scenario %>%
  mutate(omFlux = cumulativeTotalOm - lag(cumulativeTotalOm),
         fluxDirection = ifelse(omFlux>0, "positive", "negative")) %>%
  filter(complete.cases(.))

fluxPlot <- ggplot(data=scenarioFlux, aes(x=years, y=omFlux)) +
  geom_hline(aes(yintercept=0), lwd=1.25) +
  geom_step() +
  geom_point(aes(color=fluxDirection)) +
  ggtitle("Net OM Flux") +
  theme(legend.title=element_blank(),
        legend.position = "top") +
  ylab(expression(paste("Net Flux (g Organic Matter cm"^"-2",")", sep="")))

grid.arrange(plot2, fluxPlot, ncol=1)

```

## Animate Depth Profile Formation

``` {r animate cohorts code }

animateCohorts <- function(cohorts, scenario) {
  # First reshape the mass cohorts so that they're in long form
  # Calculate cumulative and additive volume 
  # Second reshape the scenario, so that we have MSL and MHW to MLW
  # If MHHW and MHHWS, do the same
  
  # gganimate stuff
  
  # return gif
  
}

```


## Animate Biomass Response to Sea-Level Change

## Animate Marsh Elevation Relative to Mean Sea-Level


But what if we have a more complicated ecology at our wetlands and we want to represent different species?

``` {r multispecies biomass function }

totalBmassElevMultiSpeciesResponse <- function(z, bMax, zVegMax, zVegMin, zVegPeak, bgTurnover, 
                                               rootShoot, rootDepth, speciesCode, presenceAbsence=c(1,1)) {
  
  # Add some stops to make sure that all the input vectors are the same length 
  if (all(is.na(zVegPeak))) {
   inputsList <- list(bMax, zVegMax, zVegMin, bgTurnover, rootShoot, rootDepth) 
  } else {
    inputsList <- list(bMax, zVegMax, zVegMin, zVegPeak, bgTurnover, rootShoot, rootDepth) 
  }
  inputLengths <- mapply(length, x = inputsList)
  if (min(inputLengths)!=max(inputLengths)) { stop("inputs are not all the same length") }
  
  if (all(is.na(speciesCode))) {
    speciesCode <- as.character(1:length(bMax))
  }
  
  # Set up a dataframe with the
  agbBySpecies <- data.frame(agb = rep(NA, length(bMax)*length(z)),
                             elevation = rep(NA, length(bMax)*length(z)),
                             speciesCode = rep(NA, length(bMax)*length(z)))
  
  # iterate through all species
  rowStart = 1
  for (i in 1:length(bMax)) {
    agbBySpecies$agb[rowStart:(i*length(z))] <- predictedBiomass(z=z, bMax=bMax[i], zVegMax=zVegMax[i],
                                                  zVegMin=zVegMin[i], zVegPeak=zVegPeak[i])
    agbBySpecies$elevation[rowStart:(i*length(z))] <- z
    agbBySpecies$speciesCode[rowStart:(i*length(z))] <- rep(speciesCode[i], length(z))
    
    rowStart <- (i*length(z))+1
  }
  
  # Get totals
  totalBmass <- agbBySpecies %>%
    group_by(elevation) %>%
    summarise(totalAgb = sum(agb))
  
  # Calculate ratio of total to species agb
  ratioOfSpToTotal <- agbBySpecies %>%
    left_join(totalBmass) %>%
    mutate(bmassRatio = agb/totalAgb) %>%
    # filter(complete.cases(.)) %>%
    select(elevation, speciesCode, bmassRatio) %>% 
    spread(key=speciesCode, value=bmassRatio) %>%
    select(-elevation)
  
  ratioOfSpToTotal[is.na(ratioOfSpToTotal)] <- 0
  
  # weighted average bgTurnover, rootShoot, and rootDepth
  weightedBgTurnover <- rowSums(t(bgTurnover * t(ratioOfSpToTotal)))
  weightedRootShoot <- rowSums(t(rootShoot * t(ratioOfSpToTotal)))
  weightedRootDepth <- rowSums(t(rootDepth * t(ratioOfSpToTotal))) 
  
  totalBmass <- totalBmass %>%
    mutate(totalBgb = bgbFromAgb(agb=totalAgb,
                                 rootShoot = weightedRootShoot))
  return(list(totalBmass, agbBySpecies, weightedRootDepth))
} 

```

Here's an example of two different plants, with similar above ground biomass, and root shoot ratio, but a higher marsh that has more turnover than the lower marsh.

``` {r multispecies biomass function example, fig.height = 3.54, fig.width=6}


# Code that could be part of the scenario developer program
  # Add species presence or absence if people want to simulate new species arriving
  # if (!missing(presenceAbsence)) {
  #   if (nrow(presenceAbsence) == nYearsOfSim) {
  #     if (ncol(presenceAbsence) == length(speciesCode)) {
  #       scenario[1:nYearsOfSim, speciesNames] <- presenceAbsence
  #     } else {
  #       stop("Species names are not defined.")
  #     }
  #   } else {
  #     stop("Species presence absence input is incorrect. Enter a dataframe equal in length equal to the number of years in the scenario.")
  #   }
  # } 

# Simple multiplications
bgbFromAgb <- function(agb, rootShoot) {agb*rootShoot} 

dynamicRootBiomass <- function(z, bMax, zVegMax, zVegMin, zVegPeak, bgTurnover, rootShoot) {
  agb <- predictedBiomass(z=z, bMax=bMax, zVegMax=zVegMax, zVegMin=zVegMin, zVegPeak=zVegPeak)
  bgb <- bgbFromAgb(agb=agb, rootShoot=rootShoot)
  return(bgb)
}

# Multiple Species Response
bMax=c(2500, 2200)
zVegMax=c(3,4)
zVegMin=c(-1,1)
zVegPeak=c(2,3.5)
bgTurnover=c(1,2.5)
rootShoot=c(2,2)
rootDepth=c(0.5,1)
speciesCode=c("SPAL","SPPA")
z <- seq(-1.1,4.1, by=0.1)

twoSpeciesExample <- totalBmassElevMultiSpeciesResponse(z=z, bMax=bMax, zVegMax=zVegMax, zVegMin=zVegMin,
                                                        zVegPeak=zVegPeak, bgTurnover=bgTurnover,
                                                        rootShoot=rootShoot, rootDepth=rootDepth,
                                                        speciesCode=speciesCode)

agbSpecies <- twoSpeciesExample[[2]] %>%
  mutate(biomassType=paste(speciesCode, " agb", sep="")) %>%
  rename(biomass = agb)

agbBgbDisplay <- twoSpeciesExample[[1]] %>%
  mutate(totalBgb = -totalBgb) %>% 
  gather(key = biomassType, value = biomass, -elevation) %>%
  bind_rows(agbSpecies)

ggplot(agbBgbDisplay, aes(x=elevation, y=biomass)) +
  geom_line(aes(color=biomassType), lwd=1.25)

```
